### 2023-03-29 ARB/standards/work-in-progress/document/tex/boilerplate/about.tex
add tex subdir

\def\tightlist{}
\chapter*{Introduction}
\section*{History of the International \rexx{} Language Symposium}
In 1990, Cathie Dager of SLAC\footnote{Stanford Linear Accelerator Center, since 2008 SLAC National Accelerator Laboratory} convened the organizing committee for the first independent \rexx{}\footnote{Cowlishaw, M. F., \textbf{The REXX Language} (second edition), ISBN 0-13-780651-5, Prentice-Hall, 1990.} Symposium for Developers and Users. SLAC continued to organize this annual event until the middle of the 1990's when the \rexx{}LA took over that responsibility. Symposia have been held annually since 1990.
\section*{About RexxLA}
During the 1993 Symposium in La Jolla, California, plans for a \rexx{} User Group materialized. The \rexx{} Language Association (\rexx{}LA), as it was called, is an independent, non-profit organization dedicated to promoting the use and understanding of the \rexx{} programming language. \rexx{}LA manages several open source implementations of \rexx{}.
\section*{The selection procedure}Presentation proposals are sollicited yearly using a CFP\footnote{Call For Papers.} procedure, after which the RexxLA symposium comittee reviews them and votes which presentations are selected for the symposium. The presentations are peer reviewed before being presented. Presenters are not compensated for their presentations.
\input{symposium}