### 2023-03-29 ARB/standards/work-in-progress/document/tex/boilerplate/acknowledgements.tex

\small
\chapter*{Acknowledgments}
\begin{multicols*}{2}
RexxLA thanks everybody who has been involved with the \rexx{} symposia over more than 30 years. Mike Cowlishaw for being the Father of \rexx{}, Rick McGuire for being the driving force behind all IBM implementations of \rexx{}, Chip Davis who is Past-President for Life, Mark Hessling for maintaining Regina and building our wonderful symposium application and web site (which enable this publication). Simon Nash for being the architect of Oryx, which became Object Rexx. Jon Wolfers for all his work in converting the website to the database-driven system, and being our treasurer. Gil Barmwater for being our vice-president and treasurer for many years. Rony Flatscher for his teaching and enthousiasm, and for connecting ooRexx to the Java Virtual Machine and its libraries. All the presenters for investing their time to deliver the message. Terry Fuller for being the current vice-president and his help during the symposia. Lee Peedin for his work as President and Vice President for many years. Cathie Dager for being a \rexx{} ``Spark Plug'', Pam Taylor for all her work for RexxLA. Virgil Hein and Matthew Emmons for being our representatives within IBM and assisting us through the open sourcing of the IBM products. We thank IBM for graciously donating two of their \rexx{} implementations, Object Rexx and NetRexx, to RexxLA and the open source community. We thank Per-Olov Jonsson for financing and running our automated test facility - and for all his work on that very important environment. David Ashley his contributions in the early days of ``Open'' Object Rexx. Marc Remes for saving NetRexx from the Java Platform Modules System. Erich Steinbock and Jean-Louis Faucher for their contributions (and future contributions) to ooRexx. 
\\[0.5cm]
RexxLA remembers our departed, Kurt Maerker, Brian Marks, Les Koehler, Mark Miesfeld, Oliver Sims and Kermit Kiser, for all their work and pleasant company.
\end{multicols*}