### 2023-03-29 ARB/standards/work-in-progress/document/tex/boilerplate/bookmeta.tex

\section*{Publication Data}
\textcopyright  Copyright The Rexx Language Association, \splice{nrc -noverbose -exec ../boilerplate/TexYear}
%\\

All original material in this publication is published under the
Creative Commons - Share Alike 3.0 License as stated at
\url{https://creativecommons.org/licenses/by-nc-sa/3.0/us/legalcode}.\\[0.5cm]

A publication of \textbf{RexxLA Press}\\[0.5cm]
The responsible publisher of this edition is identified as \emph{IBizz
IT Services and Consultancy}, Amsteldijk 14, 1074 HR Amsterdam, a
registered company governed by the laws of the Kingdom of The
Netherlands.\\[1cm]
The RexxLA Symposium Series is registered under ISSN 1534-8954\\
The \edition{} edition is registered under ISBN \isbn \\[1cm]
\psset{unit=1in}
\begin{pspicture}(3.5,1in)
  \psbarcode{1534-8954}{includetext guardwhitespace}{issn}
\end{pspicture}
\begin{pspicture}(3.5,1in)
  \psbarcode{\isbn}{includetext guardwhitespace}{isbn}
\end{pspicture}
%\section*{Publication History}
\\[0.5cm]
\begin{description}
\item[2023-03-14] First printing
\end{description}
\newpage
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: