### 2023-05-02 ARB/standards/work-in-progress/document/tex/standard/built-in_functions.tex

\hypertarget{built-in-functions}{%
\chapter{Built-in functions}\label{built-in-functions}}

\hypertarget{notation}{%
\section{Notation}\label{notation}}

The built-in functions are defined mainly through code. The code refers
to state variables. This is solely a notation used in this standard.

The code refers to functions with names that start with `Config\_';
these are the functions described in section nnn.

The code is specified as an external routine that produces a result from
the values \#Bif (which is the name of the built-in function),
\#Bif\_Arg.0 (the number of arguments), \#Bif\_Arg.i and
\#Bif\_ArgExists.i (which are the argument data.)

The value of \#Level is the value for the clause which invoked the
built-in function.

The code either returns the result of the built-in or exits with an
indication of a condition that the invocation of the built-in raises.

The code below uses built-in functions. Such a use invokes another use
of this code with a new value of \#Level. On these invocations, the
CheckArgs function is not relevant.

Numeric settings as follows are used in the code. When an argument is
being checked as a number by `NUM' or `WHOLENUM' the settings are those
current in the caller. When an argument is being checked as an integer
by an item containing `WHOLE' the settings are those for the particular
built-in function. Elsewhere the settings have sufficient numeric digits
to avoid values which would require exponential notation.

\hypertarget{routines-used-by-built-in-functions}{%
\section{Routines used by built-in
functions}\label{routines-used-by-built-in-functions}}

The routine CheckArgs is concerned with checking the arguments to the
built-in. The routines Time2Date and Leap are for date calculations.
ReRadix is used for radix conversion. The routine Raise raises a
condition and does not return.

\hypertarget{argument-checking}{%
\subsection{Argument checking}\label{argument-checking}}

/* Check arguments. Some further checks will be made in particular
built-ins.\emph{/ /} The argument to CheckArgs is a checklist for the
allowable arguments. */

/* NUM, WHOLENUM and WHOLE have a side-effect, `normalizing' the number.
*/

/* Calls to raise syntax conditions will not return. */

CheckArgs: CheckList = arg(1) /* This refers to the argument of
CheckArgs. */

/* Move the checklist information from a string to individual variables
\emph{/ ArgType. = '\,'\,' ArgPos = 0 /} To count arguments \emph{/
MinArgs = 0 do j = 1 to length (CheckList) ArgPos = ArgPos+1 /} Count
the required arguments. \emph{/ if substr(CheckList,j,1) == `r' then
MinArgs = MinArgs + 1 /} Collect type information. \emph{/ do while j
\textless{} length(CheckList) j=eajae1 t = substr(CheckList,j,1) if t=='
' then leave ArgType.ArgPos = ArgType.ArgPos \textbar\textbar{} t end /}
A single space delimits parts. */ end j MaxArgs = ArgPos

/* Check the number of arguments to the built-in, in this instance. */
NumArgs = \#Bif Arg.0

if NumArgs \textless{} MinArgs then call Raise 40.3, MinArgs

if NumArgs \textgreater{} MaxArgs then call Raise 40.4, MaxArgs

/* Check the type(s) of the arguments to the built-in. */ do ArgPos = 1
to NumArgs if \#Bif ArgExists.ArgPos then call CheckType else if ArgPos
\textless= MinArgs then call Raise 40.5, ArgPos end ArgPos

/* No errors found by CheckArgs. */

return CheckType: Value = \#Bif Arg.ArgPos Type = ArgType.ArgPos select
when Type == `ANY' then nop /* Any string \emph{/ when Type == `NUM'
then do /} Any number */

/* This check is made with the caller's digits setting. \emph{/ if
\Cdatatype(Value, `N') then if \#DatatypeResult==`E' then call Raise
40.9, ArgPos, Value else call Raise 40.11, ArgPos, Value \#Bif
Arg.ArgPos=\#DatatypeResult /} Update argument copy. */ end

when Type == `WHOLE' then do /* Whole number \emph{/ /} This check is
made with digits setting for the built-in. */ if \Edatatype(Value,`W')
then call Raise 40.12, ArgPos, Value \#Bif\_ Arg.ArgPos=\#DatatypeResult
end

when Type == `WHOLE\textgreater=0' then do /* Non-negative whole number
*/ if \Edatatype(Value,`W') then call Raise 40.12, ArgPos, Value if
\#DatatypeResult \textless{} 0 then call Raise 40.13, ArgPos, Value
\#Bif\_ Arg.ArgPos=\#DatatypeResult end

when Type == `WHOLE\textgreater0O' then do /* Positive whole number */
if \Edatatype(Value,`W') then call Raise 40.12, ArgPos, Value if
\#DatatypeResult \textless= 0 then call Raise 40.14, ArgPos, Value
\#Bif\_ Arg.ArgPos=\#DatatypeResult end

when Type == `WHOLENUM' then do /* D2X type whole number \emph{/ /} This
check is made with digits setting of the caller. */ if
\Cdatatype(Value,`W') then call Raise 40.12, ArgPos, Value \#Bif\_
Arg.ArgPos=\#DatatypeResult end

when Type == `WHOLENUM\textgreater=0' then do /* D2X Non-negative whole
number */ if \Cdatatype(Value,`W') then call Raise 40.12, ArgPos, Value
if \#DatatypeResult \textless{} 0 then call Raise 40.13, ArgPos, Value
\#Bif\_ Arg.ArgPos=\#DatatypeResult end

when Type == `0 90' then do /* Errortext */

if \Edatatype(Value,`N') then call Raise 40.11, ArgPos, Value

Value=\#DatatypeResult

\#Bif\_ Arg.ArgPos=Value

Major=Value \% 1

Minor=Value - Major

if Major \textless{} 0 \textbar{} Major \textgreater{} 90 \textbar{}
Minor \textgreater{} .9 \textbar{} pos(`E',Value)\textgreater0 then call
Raise 40.17, Value /* ArgPos will be 1 */

end

when Type == `PAD' then do /* Single character, usually a pad. */

if length(Value) = 1 then call Raise 40.23, ArgPos, Value end

when Type == `HEX' then /* Hexadecimal string \emph{/ if
\datatype (Value, `X') then call Raise 40.25, Value /} ArgPos will be 1
\emph{/ when Type == `BIN' then /} Binary string \emph{/ if
\datatype(Value,`B') then call Raise 40.24, Value /} ArgPos will be 1
\emph{/ when Type == `SYM' then /} Symbol */

if \datatype (Value, `S') then call Raise 40.26, Value /* ArgPos will be
1 */

when Type == `STREAM' then do call Config Stream Qualify Value if left
(\#Response, 1) == `B' then call Raise 40.27, Value /* ArgPos will be 1
*/ end

when Type = `ACEFILNOR' then do /* Trace \emph{/ Val = Value /} Allow
`?' alone */

if val == `?' then do /* Allow leading `?' */ if left(Val,1) == `?' then
Val = substr(Val,2)

if pos(translate(left(Val, 1)), `ACEFILNOR') = 0 then call Raise 40.28,
ArgPos, Type, Val

end end otherwise do /* Options \emph{/ /} The checklist item is a list
of allowed characters */ if Value == '\,' then

call Raise 40.21, ArgPos \#Bif Arg.ArgPos = translate(left(Value, 1)) if
pos(\#Bif Arg.ArgPos, Type) = 0 then call Raise 40.28, ArgPos, Type,
Value end

end /* Select */ return

Cdatatype: /* This check is made with the digits setting of the caller.
\emph{/ /} \#DatatypeResult will be set by use of datatype() */

numeric digits \#Digits.\#Level

numeric form value \#Form.\#Level

return datatype(arg(1), arg(2))

Edatatype: /* This check is made with digits setting for the particular
built-in. \emph{/ /} \#DatatypeResult will be set by use of datatype()
*/

numeric digits \#Bif Digits.\#Bif

numeric form scientific

return datatype(arg(1),arg(2))

10.2.2 Date calculations Time2Date: if arg(1) \textless{} 0 then call
Raise 40.18 if arg(1) \textgreater= 315537897600000000 then call Raise
40.18 return Time2Date2 (arg(1))

Time: procedure /* This routine is essentially the code from the
standard, put in stand-alone form. The only `tricky bit' is that there
is no Rexx way for it to fail with the same error codes as a ``real''
implementation would. It can however give a SYNTAX error, albeit not the
desirable one. This causing of an error is done by returning with no
value. Since the routine will have been called as a function, this
produces an error. */

/* Backslash is avoided as some systems don't handle that negation sign.
\emph{/ if argQ)\textgreater3 then returm numeric digits 18 if
arg(1,`E') then if pos(translate(left(arg(1),1)),``CEHLMNRS'')=0 then
return /} (The standard would also allow `O' but what this code is
running on would not.) \emph{/ if arg(3,`E') then if
pos(translate(left(arg(3),1)),``CHLMNS'')=0 then return /} Tf the third
argument is given then the second is mandatory. \emph{/ if arg(3,`E') \&
arg(2,`E')=0 then returm /} Default the first argument. \emph{/ if
arg(1,`E') then Option = translate(left(arg(1),1)) else Option = `N'' /}
Tf there is no second argument, the current time is returned. \emph{/ if
arg(2,`E') = 0 then if arg(1,`E') then return ``TIME'(arg(1)) else
return `TIME'() /} One cannot convert to elapsed times. */ if
pos(Option, `ERO') \textgreater{} 0 then returm InValue = arg(2) if
arg(3,`E') then InOption = arg(3) else InOption = `N' HH =0 MM = 0 SS =0
HourAdjust = 0 select

when InOption == `C' then do parse var InValue HH `:'. +1 MM +2 XX if HH
= 12 then HH =0 if XX == `pm' then HourAdjust = 12

end when InOption == `H' then HH = InValue

when InOption == `L'' \textbar{} InOption == `N' then parse var InValue
HH `:' MM `:' SS when InOption == `M' then MM = InValue otherwise SS =
InValue end if datatype(HH,`W')=0 \textbar{} datatype(MM,`W')=0
\textbar{} datatype(SS,`N')=0 then returm HH = HH + HourAdjust /*
Convert to microseconds \emph{/ Micro = trunc((((HH } 60) + MM) * 60 +
SS) * 1000000) /* There is no special message for time-out-of-range; the
bad-format message is used. \emph{/ if Micro\textless0O \textbar{} Micro
\textgreater{} 24}3600* 1000000 then returm /* Reconvert to further
check the original. */ if TimeFormat(Micro,InOption) == InValue then
return TimeFormat(Micro, Option) return

TimeFormat: procedure /* Convert from microseconds to given format.
\emph{/ /} The day will be irrelevant; actually it will be the first day
possible. */ x = Time2Date2(arg(1)) parse value x with Year Month Day
Hour Minute Second Microsecond Base Days select when arg(2) == `C' then
select when Hour\textgreater12 then return
Hour-12':`right(Minute,2,'0'')'pm' when Hour=12 then return
`12:'right(Minute,2,'0')`pm' when Hour\textgreater0 then return
Hour':`right(Minute,2,'0')`am' when Hour=0 then return
`12:'right(Minute,2,'0')`am'

when arg(2) == `H' then return Hour when arg(2) == `L' then return
right(Hour,?2,`0')`:'right(Minute,2,'0')`:'right(Second,2,'0'),
\textbar\textbar{} `.'right(Microsecond,6,'0') when arg(2) == `M' then
return 60\emph{Hour+Minute when arg(2) == `N' then return
right(Hour,?2,`0')`:'right(Minute,2,'0')`:'right(Second,2,'0') otherwise
/} arg(2) ==`S' \emph{/ return 3600}Hour+60* Minute+Second end

Time2Date2: Procedure /* Convert a timestamp to a date. Argument is a
timestamp (the number of microseconds relative to 0001 01 01
00:00:00.000000) Returns a date in the form: year month day hour minute
second microsecond base days */

/* Argument is relative to the virtual date 0001 01 01 00:00:00.000000
*/ Time = arg(1)

Second = Time \% 1000000 \} Microsecond = Time // 1000000 Minute =
Second \% 60 ; Second = Second // 60 Hour = Minute \% 60 ; Minute =
Minute // 60 Day = Hour \% 24 ;\} Hour = Hour // 24

/* At this point, the days are the days since the 0001 base date. */
BaseDays = Day Day = Day + 1

/* Compute either the fitting year, or some year not too far earlier.
Compute the number of days left on the first of January of this year. */

Year = Day \% 366 Day = Day - (Year*365 + Year\%4 - Year\%100 +
Year\%400) Year = Year +1

/* Now if the number of days left is larger than the number of days in
the year we computed, increment the year, and decrement the number of
days accordingly. */ do while Day \textgreater{} (365 + Leap (Year) )

Day = Day - (365 + Leap(Year) )

Year = Year + 1 end

/* At this point, the days left pertain to this year. */ YearDays = Day

/* Now step through the months, increment the number of the month, and
decrement the number of days accordingly (taking into consideration that
in a leap year February has 29 days), until further reducing the number
of days and incrementing the month would lead to a negative number of
days */ Days = `31 28 31 30 31 30 31 31 30 31 30 31' do Month = 1 to
words (Days)

ThisMonth = Word(Days, Month) + (Month = 2) * Leap (Year)

if Day \textless= ThisMonth then leave

Day = Day - ThisMonth end

return Year Month Day Hour Minute Second Microsecond BaseDays YearDays

Leap: procedure /* Return 1 if the year given as argument is a leap
year, or 0 otherwise. */ return (arg(1)//4 = 0) \& ((arg(1)//100
\textless\textgreater{} 0) \textbar{} (arg(1)//400 = 0))

10.2.1. Radix conversion

ReRadix: /* Converts Arg(1) from radix Arg(2) to radix Arg(3) */

procedure

Subj ect=arg(1)

FromRadix=arg (2)

ToRadix=arg (3)

/* Radix range is 2-16. Conversion is via decimal */

Integer=0

do j=1 to length (Subject) /* Individual digits have already been
checked for range. \emph{/ Integer=Integer}FromRadix+pos (substr
(Subject,j,1),`0123456789ABCDEF')-1 end

rete

do while Integer\textgreater0

r= substr(`0123456789ABCDEF',1 + Integer // ToRadix, 1)
\textbar\textbar{} r Integer = Integer \% ToRadix end

/* When between 2 and 16, there is no zero suppression. \emph{/ if
FromRadix = 2 \& ToRadix = 16 then r=eright(r, (length(Subject)+3) \% 4,
`0') else if FromRadix = 16 \& ToRadix = 2 then reright(r,
length(Subject) } 4, `0') return r

\hypertarget{raising-the-syntax-condition}{%
\subsection{Raising the SYNTAX
condition}\label{raising-the-syntax-condition}}

Raise:

/* These 40.nn messages always include the built-in name as an
insert.\emph{/ call \#Raise `SYNTAX', arg(1), \#Bif, arg(2), arg(3),
arg(4) /} \#Raise does not return. */

\hypertarget{character-built-in-functions}{%
\section{Character built-in
functions}\label{character-built-in-functions}}

These functions process characters or words in strings. Character
positions are numbered from one at the left. Words are delimited by
blanks and their equivalents, word positions are counted from one at the
left.

\hypertarget{abbrev}{%
\subsection{ABBREV}\label{abbrev}}

ABBREV returns `1' if the second argument is equal to the leading
characters of the first and the length of the second argument is not
less than the third argument.

call CheckArgs `rANY rANY oWHOLE\textgreater=0'

Subject \#Bif Arg.1

Subj \#Bif Arg.2

if \#Bif\_ArgExists.3 then Length = \#Bif Arg.3 else Length = length
(Subj)

Condl = length(Subject) \textgreater= length(Subj) Cond2 = length(Subj)
\textgreater= Length Cond3 = substr(Subject, 1, length(Subj)) == Subj

return Condl \& Cond2 \& Cond3

\hypertarget{center}{%
\subsection{CENTER}\label{center}}

CENTER returns a string with the first argument centered in it. The
length of the result is the second argument and the third argument
specifies the character to be used for padding.

call CheckArgs `rANY rWHOLE\textgreater=0 oPAD'

String = \#Bif Arg.1

Length = \#Bif Arg.2

if \#Bif\_ArgExists.3 then Pad = \#Bif Arg.3 else Pad = ' '!

Trim = length(String) - Length

if Trim \textgreater{} 0 then return substr(String, Trim \% 2 + 1,
Length)

return overlay(String, copies(Pad, Length), -Trim \% 2 + 1)

\hypertarget{centre}{%
\subsection{CENTRE}\label{centre}}

This is an alternative spelling for the CENTER built-in function.

\hypertarget{changestr}{%
\subsection{CHANGESTR}\label{changestr}}

CHANGESTR replaces all occurrences of the first argument within the
second argument, replacing them with the third argument.

call CheckArgs ```rANY rANY rANY'

Output = '\,' Position = 1 do forever FoundPos = pos(\#Bif Arg.1, \#Bif
Arg.2, Position) if FoundPos = 0 then leave Output = Output
\textbar\textbar{} substr(\#Bif\_ Arg.2, Position, FoundPos - Position),
\textbar\textbar{} \#Bif Arg.3 Position = FoundPos + length(\#Bif Arg.1)
end return Output \textbar\textbar{} substr(\#Bif Arg.2, Position)

\hypertarget{compare}{%
\subsection{COMPARE}\label{compare}}

COMPARE returns `0' if the first and second arguments have the same
value. Otherwise, the result is the position of the first character that
is not the same in both strings.

call CheckArgs `rANY rANY oPAD'

Strl = \#Bif\_Arg.1

Str2 = \#Bif\_Arg.2

if \#Bif\_ArgExists.3 then Pad else Pad

\#Bif Arg.3

/* Compare the strings from left to right one character at a time */ if
length(Str1) \textgreater{} length(Str2) then do

Length = length(Str1)

Str2=left (Str2, Length, Pad)

end else do

Length = length(Str2)

Strl=left (Str1, Length, Pad)

end

do i= 1 to Length if substr(Stri, i, 1) == substr(Str2, i, 1) then
return i end

return 0

\hypertarget{copies}{%
\subsection{COPIES}\label{copies}}

COPIES returns concatenated copies of the first argument. The second
argument is the number of copies.

call CheckArgs ```rANY rWHOLE\textgreater=0' Output = '\,' do \#Bif
Arg.2 Output = Output \textbar\textbar{} \#Bif\_Arg.1 end

return Output

\hypertarget{countstr}{%
\subsection{COUNTSTR}\label{countstr}}

COUNTSTR counts the appearances of the first argument in the second
argument.

call CheckArgs ```rANY rANY'

Output = 0

Position = pos (\#Bif Arg.1,\#Bif Arg.2)

do while Position \textgreater{} 0 Output = Output + 1 Position =
pos(\#Bif Arg.1, \#Bif Arg.2, Position + length(\#Bif Arg.1)) end

return Output

\hypertarget{datatype}{%
\subsection{DATATYPE}\label{datatype}}

DATATYPE tests for characteristics of the first argument. The second
argument specifies the particular test.

call CheckArgs `rANY oABLMNSUWX'

/* As well as returning the type, the value for a `NUM' is set in
\#DatatypeResult. This is a convenience when DATATYPE is used

by CHECKARGS. */

String = \#Bif Arg.1

/* If no second argument, DATATYPE checks whether the first is a number.
*/ if \#Bif\_ArgExists.2 then return DtypeOne()

Type = \#Bif\_ Arg.2 /* Null strings are a special case. */

if String == '\,' then do if Type == ``X'' then return 1 if Type ==
``B'' then return 1 return 0 end

/* Several of the options are shorthands for VERIFY \emph{/
azl=``abcdefghijklmnopqrstuvwxyz'' AZU= ``ABCDEFGHIJKLMNOPQRSTUVWXYZ''
DO09=``0123456789'' if Type == ``A'' then return
verify(String,az1\textbar\textbar AzU\textbar\textbar D09)=0 if Type ==
``B'' then do /} Check blanks in allowed places. \emph{/ if pos (left
(String,1),\#Al1Blanks)\textgreater0 then return 0 if pos (right
(String,1),\#Al1Blanks)\textgreater0 then return 0 BinaryDigits=0 do j =
length(String) by -1 tol ce = substr(String,j,1) if
pos(c,\#Al11Blanks)\textgreater0 then do /} Blanks need four
BinaryDigits to the right of them. */ if BinaryDigits//4 = 0 then return
0 end else do if verify(c,``01'') = 0 then return 0 BinaryDigits =
BinaryDigits + 1

end end j

return 1

end /* B */ if Type == ``L'' then return (verify (String,azl1)=0) if
Type == ``M'' then return (verify (String, azl1\textbar\textbar AZU)=0)
if Type == ``N'' then return (datatype (String) ==``NUM'') if Type ==
``S'' then return(symbol (String) ==`BAD') if Type == ``U'' then return
(verify (String, AZU)=0) if Type == ``W'' then do

/* It may not be a number. \emph{/ if DtypeOne(String) == `CHAR' then
return `0' /} It can be ``Whole'' even if originally in exponential
notation, provided it can be written as non-exponential. \emph{/ if
pos(`E',\#DatatypeResult)\textgreater0 then return `0' /} It won't be
``Whole'' if there is a non-zero after the decimal point. */
InFraction=`0'

do j = 1 to length (String) ce = substr(String,j,1) if
pos(c,`Ee')\textgreater0 then leave j if InFraction \&
pos(c,`+-')\textgreater0 then leave j if c == `.' then InFraction=`1'
else if InFraction \& a==`0' then return 0 end j /* All tests for Whole
passed. \emph{/ \#DatatypeResult = \#DatatypeResult \% 1 return 1 end /}
W \emph{/ /} Type will be ``x'' \emph{/ if pos (left
(String,1),\#Al1Blanks)\textgreater0 then return 0 if pos (right
(String,1),\#Al1Blanks)\textgreater0 then return 0 HexDigits=0 do
j=length(String) by -1 to 1 c=substr (String,j,1) if
pos(c,\#Al11Blanks)\textgreater0 then do /} Blanks need a pair of
HexDigits to the right of them. \emph{/ if HexDigits//2 = 0 then return
0 end else do if verify(c,``abcdefABCDEF''D09) = 0 then return 0
HexDigits=HexDigits+1 end end return 1 /} end X */

DtypeOne: /* See section nnn for the syntax of a number. \emph{/
\#DatatypeResult = `S' /} If not syntactically a number \emph{/ Residue
= strip(String) /} Blanks are allowed at both ends. */ if Residue ==
'\,' then return ``CHAR'' Sign = '\,' if left(Residue,1) == `+'
\textbar{} left(Residue,1) == `-' then do

Sign = left(Residue, 1) Residue = strip(substr(Residue,2),`L') /* Blanks
after sign */

end if Residue == '\,' then return ``CHAR'' /* Now testing Number,
section nnn */ if left(Residue,1) == `.' then do Residue = substr
(Residue, 2) Before = '\,' After = DigitRun() if After == '\,' then
return ``CHAR'' end else do Before = DigitRun() if Before == '\,' then
return ``CHAR'' if left(Residue,1) == `.' then do Residue = substr
(Residue, 2) After = DigitRun() end end Exponent = 0 if Residue == '\,'
then do if left(Residue, 1) == `e' \& left(Residue, 1) == `E' then

return ``CHAR'' Residue = substr (Residue, 2)

if Residue == '\,' then return ``CHAR''

Esign = '\,'

if left(Residue, 1) == `+' \textbar{} left(Residue, 1) == `-' then do
Esign = left(Residue, 1) Residue = substr (Residue, 2) if Residue ==
'\,' then return ``CHAR'' end

Exponent = DigitRun()

if Exponent == '\,' then return ``CHAR''

Exponent = Esign \textbar\textbar{} Exponent

end

if Residue == '\,' then return ``CHAR''

102 /\emph{DATATYPE tests for exponent out of range. }/

\#DatatypeResult = `E' /* If exponent out of range */ Before =
strip(Before,`L',`0') if Before == '\,' then Before = `0'

Exponent = Exponent + length(Before) -1 /* For SCIENTIFIC */

/* ``Engineering notation causes powers of ten to expressed as a
multiple of 3 - the integer part may therefore range from 1 through
9910.'' \emph{/ g=1 if \#Form.\#Level == `E' then do /} Adjustment to
make exponent a multiple of 3 */ g = Exponent//3 if g \textless{} 0 then
g = g + Exponent = Exponent - end

/* Check on the exponent. */ if Exponent \textgreater{} \#Limit
ExponentDigits then return ``CHAR'' if -\#Limit ExponentDigits
\textgreater{} Exponent then return ``CHAR''

/* Format to the numeric setting of the caller of DATATYPE */ numeric
digits \#Digits.\#Level

numeric form value \#Form.\#Level

\#DatatypeResult = 0 + \#Bif\_Arg.1

return ``NUM''

DigitRun: Outcome = '\,' do while Residue == '\,' if pos(left (Residue,
1), `0123456789') = 0 then leave

Outcome = Outcome \textbar\textbar{} left(Residue, 1) Residue =
substr(Residue, 2) end

return Outcome

\hypertarget{delstr}{%
\subsection{DELSTR}\label{delstr}}

DELSTR deletes the sub-string of the first argument which begins at the
position given by the second argument. The third argument is the length
of the deletion.

call CheckArgs `rANY rWHOLE\textgreater0 oWHOLE\textgreater=0' String
\#Bif Arg.1

Num \#Bif Arg.2 if \#Bif\_ArgExists.3 then Len = \#Bif\_Arg.3

if Num \textgreater{} length(String) then return String

Output = substr(String, 1, Num - 1) if \#Bif\_ArgExists.3 then if Num +
Len \textless= length(String) then Output = Output \textbar\textbar{}
substr(String, Num + Len) return Output

\hypertarget{delword}{%
\subsection{DELWORD}\label{delword}}

DELWORD deletes words from the first argument. The second argument
specifies position of the first word to be deleted and the third
argument specifies the number of words.

call CheckArgs `rANY rWHOLE\textgreater0 oWHOLE\textgreater=0' String
\#Bif Arg.1

Num \#Bif Arg.2 if \#Bif\_ArgExists.3 then Len = \#Bif\_Arg.3

if Num \textgreater{} words(String) then return String

EndLeft = wordindex(String, Num) - 1 Output = left(String, EndLeft) if
\#Bif\_ArgExists.3 then do BeginRight = wordindex(String, Num + Len)

if BeginRight\textgreater0 then Output =

end return Output

\hypertarget{insert}{%
\subsection{INSERT}\label{insert}}

Output \textbar\textbar{} substr(String, BeginRight)

INSERT insets the first argument into the second. The third argument
gives the position of the character before the insert and the fourth
gives the length of the insert. The fifth is the padding character.

call CheckArgs

New \#Bif\_ Arg.1 Target \#Bif Arg.2 if \#Bif\_ArgExists.3

then else then else then else

Num Num Length = \#Bif Arg.4 Length = length (New) Pad \#Bif Arg.5 Pad
ro

\#Bif Arg.3 0

if \#Bif\_ArgExists.4

if \#Bif\_ArgExists.5

return left(Target, Num, Pad) left (New, Length, Pad), substr(Target,
Num + 1)

``

\hypertarget{lastpos}{%
\subsection{LASTPOS}\label{lastpos}}

/* To left of insert /* New string inserted \emph{/ /} To right of
insert

`rANY rANY oWHOLE\textgreater=0 oWHOLE\textgreater=0 oPAD'

\emph{/ }/

LASTPOS returns the position of the last occurrence of the first
argument within the second. The third

argument is a starting position for the search.

call CheckArgs `rANY rANY oWHOLE\textgreater0'

Needle = \#Bif\_Arg.1 Haystack = \#Bif Arg.2 if \#Bif\_ArgExists.3 then
Start = \#Bif\_Arg.3 else Start = length(Haystack)

NeedleLength = length (Needle) if NeedleLength = 0 then return 0 Start =
Start - NeedleLength + 1 do i= Start by -1 while i \textgreater{} 0 if
substr(Haystack, i, NeedleLength) end i return 0

\hypertarget{left}{%
\subsection{LEFT}\label{left}}

LEFT returns characters that are on the left of the first argument.

length of the result and the third is the padding character. call
CheckArgs `rANY rWHOLE\textgreater=0 oPAD'

if \#Bif\_ArgExists.3 then Pad else Pad

\#Bif Arg.3

return substr(\#Bif Arg.1, 1, \#Bif\_ Arg.2, Pad)

10.1.14 LENGTH

Needle then return i

The second argument specifies the

Length returns a count of the number of characters in the argument.

call CheckArgs `rANY'

String = \#Bif Arg.1

\#Response = Config Length(String)

Length = \#Outcome

call Config Substr \#Response, 1

if \#Outcome == `E' then return Length

/* Here if argument was not a character string. call Config C2B String
call \#Raise `SYNTAX',

23.1, b2x(\#Outcome)

\emph{/ /} No return to here */

\hypertarget{overlay}{%
\subsection{OVERLAY}\label{overlay}}

OVERLAY overlays the first argument onto the second. The third argument
is the starting position of the

overlay. The fourth argument is the length of the overlay and the fifth
is the padding character.

call CheckArgs `rANY rANY oWHOLE\textgreater0 oOWHOLE\textgreater=0
oPAD'

New = \#Bif\_Arg.1

Target = \#Bif Arg.2

if \#Bif\_ArgExists.3 then Num = \#Bif\_Arg.3 else Num = 1

if \#Bif\_ArgExists.4 then Length = \#Bif\_ Arg.4 else Length = length
(New)

if \#Bif\_ArgExists.5 then Pad = \#Bif\_ Arg.5 else Pad = ' ' return
left(Target, Num - 1, Pad), /* To left of overlay \emph{/
\textbar\textbar{} left (New, Length, Pad), /} New string overlaid
\emph{/ \textbar\textbar{} substr (Target, Num + Length) /} To right of
overlay */

10.1.16 POS POS returns the position of the first argument within the
second.

call CheckArgs `rANY rANY oWHOLE\textgreater0'

Needle \#Bif Arg.1

Haystack = \#Bif Arg.2

if \#Bif\_ ArgExists.3 then Start else Start

\#Bif Arg.3 1

if length(Needle) = 0 then return 0

do i = Start to length (Haystack) +1-length (Needle) if substr(Haystack,
i, length(Needle)) == Needle then return i end i

return 0

\hypertarget{reverse}{%
\subsection{REVERSE}\label{reverse}}

REVERSE returns its argument, swapped end for end.

call CheckArgs `rANY'

String

\#Bif Arg.1 Output = '\,' do i= 1 to length (String) Output =
substr(String,i,1) \textbar\textbar{} Output end return Output

\hypertarget{right}{%
\subsection{RIGHT}\label{right}}

RIGHT returns characters that are on the right of the first argument.
The second argument specifies the

length of the result and the third is the padding character. call
CheckArgs `rANY rWHOLE\textgreater=0 oPAD'

String = \#Bif Arg.1 Length = \#Bif Arg.2 if \#Bif\_ArgExists.3 then Pad

\#Bif Arg.3 else Pad '

Trim = length(String) - Length if Trim \textgreater= 0 then return
substr(String,Trim + 1) return copies(Pad, -Trim) \textbar\textbar{}
String /* Pad string on the left */

\hypertarget{space}{%
\subsection{SPACE}\label{space}}

SPACE formats the blank-delimited words in the first argument with pad
characters between each word. The second argument is the number of pad
characters between each word and the third is the pad character.

call CheckArgs `rANY oOWHOLE\textgreater=0 oPAD'

String = \#Bif Arg.1

if \#Bif\_ArgExists.2 then Num = \#Bif\_ Arg.2 else Num = 1 if \#Bif
ArgExists.3 then Pad = \#Bif Arg.3 else Pad = ' '! Padding = copies(Pad,
Num) Output = subword(String, 1, 1) do i = 2 to words (String) Output =
Output \textbar\textbar{} Padding \textbar\textbar{} subword(String, i,
1)

end return Output

\hypertarget{strip}{%
\subsection{STRIP}\label{strip}}

STRIP removes characters from its first argument. The second argument
specifies whether the deletions are leading characters, trailing
characters or both. Each character deleted is equal to the third
argument, or equivalent to a blank if the third argument is omitted.

call CheckArgs `rANY oLTB oPAD'

String = \#Bif Arg.1 if \#Bif\_ArgExists.2 then Option = \#Bif Arg.2
else Option = `B' if \#Bif\_ ArgExists.3 then Unwanted = \#Bif\_ Arg.3
else Unwanted = \#A11Blanks\textless Index ``\#Al11Blanks'' \# ``\,''
\textgreater{}

if Option == `L' \textbar{} Option == `B' then do /* Strip leading
characters */ do while String == '\,' \& pos(left(String, 1), Unwanted)
\textgreater{} 0 String = substr(String, 2) end end

if Option == `T' \textbar{} Option == `B' then do /* Strip trailing
characters \emph{/ do while String == '\,' \& pos(right(String, 1),
Unwanted) \textgreater{} 0 String = left(String, length(String) -1) end
/} of while */ end return String

\hypertarget{substr}{%
\subsection{SUBSTR}\label{substr}}

SUBSTR returns a sub-string of the first argument. The second argument
specifies the position of the first character and the third specifies
the length of the sub-string. The fourth argument is the padding

character.

call CheckArgs `rANY rWHOLE\textgreater0 oOWHOLE\textgreater=0 oPAD'

String = \#Bif Arg.1 Num = \#Bif\_ Arg.2 if \#Bif\_ArgExists.3 then
Length = \#Bif Arg.3 else Length = max(length (String) +1-Num, 0) if
\#Bif\_ArgExists.4 then Pad = \#Bif\_ Arg.4 else Pad = ' `! Output =''
do Length \#Response Config Substr(String,Num) /* Attempt to fetch
character.*/ Character \#Outcome

Num = Num + 1 call Config Substr \#Response,1 /* Was there such a
character? */ if \#Outcome == `E' then do

/* Here if argument was not a character string. */

call Config C2B String call \#Raise `SYNTAX', 23.1, b2x(\#Outcome) /* No
return to here */

end if \#Outcome == `M' then Character = Pad Output=Output \textbar{}
\textbar{} Character end

return Output

\hypertarget{subword}{%
\subsection{SUBWORD}\label{subword}}

SUBWORD returns a sub-string of the first argument, comprised of words.
The second argument is the position in the first argument of the first
word of the sub-string. The third argument is the number of words in the
sub-string.

call CheckArgs `rANY rWHOLE\textgreater0 oWHOLE\textgreater=0'

String \#Bif Arg.1

Num \#Bif Arg.2

if \#Bif\_ ArgExists.3 then Length else Length

\#Bif\_Arg.3 length(String) /* Avoids call \emph{/ /} to WORDS() */

if Length = 0 then return '\,'

/* Find position of first included word */

Start = wordindex (String, Num)

if Start = 0 then return '\,' /* Start is beyond end */

/* Find position of first excluded word */ End = wordindex (String,
Num+Length) if End = 0 then End = length(String)+1

Output=substr (String, Start, End-Start)

/* Drop trailing blanks */

do while Output == '\,' if pos (right (Output,1),\#A11Blanks) = 0 then
leave Output = left (Output, length (Output) -1) end

return Output

\hypertarget{translate}{%
\subsection{TRANSLATE}\label{translate}}

TRANSLATE returns the characters of its first argument with each
character either unchanged or translated to another character.

call CheckArgs `rANY oANY oANY oPAD' String = \#Bif Arg.1 /* If neither
input nor output tables, uppercase. */ if \#Bif ArgExists.2 \&
\#Bif\_ArgExists.3 then do Output = '\,' do j=1 to length (String)
\#Response = Config Upper (substr (String,j,1))

Output = Output \textbar\textbar{} \#Outcome end j

return Output

end

/* The input table defaults to all characters. \emph{/ if \#Bif
ArgExists.3 then do \#Response = Config Xrange() Tablei = \#Outcome end
else Tablei = \#Bif\_Arg.3 /} The output table defaults to null \emph{/
if \#Bif\_ArgExists.2 then Tableo = \#Bif Arg.2 else Tableo = '\,' /}
The tables are made the same length */ if \#Bif\_ArgExists.4 then Pad =
\#Bif\_ Arg.4 else Pad = ' ' Tableo=left (Tableo, length (Tablei) , Pad)

107 Output='\,' do j=1 to length (String) c=substr (String,j,1)
k=pos(c,Tablei) if k=0 then Output=Output \textbar\textbar c else
Output=Output \textbar{} \textbar{} substr (Tableo,k,1) end j return
Output

\hypertarget{verify}{%
\subsection{VERIFY}\label{verify}}

VERIFY checks that its first argument contains only characters that are
in the second argument, or that it contains no characters from the
second argument; the third argument specifies which check is made. The
result is `0', or the position of the character that failed
verification. The fourth argument is a starting position for the check.

call CheckArgs `rANY rANY oMN oWHOLE\textgreater0'

String = \#Bif Arg.1

Reference = \#Bif\_Arg.2

if \#Bif\_ ArgExists.3 then Option \#Bif Arg.3 else Option 'N!

if \#Bif\_ ArgExists.4 then Start = \#Bif\_Arg.4 else Start = 1

Last = length(String) if Start \textgreater{} Last then return 0 if
Reference == '\,' then if Option == `N' then return Start else return 0

do i = Start to Last t = pos(substr(String, i, 1), Reference) if Option
== `N' then do if t = 0 then return i /* Return position of NoMatch
character. \emph{/ end else if t \textgreater{} 0 then return i /}
Return position of Matched character. */ end i return 0

\hypertarget{word}{%
\subsection{WORD}\label{word}}

WORD returns the word from the first argument at the position given by
the second argument.

call CheckArgs `rANY rwWHOLE\textgreater0'

return subword(\#Bif Arg.1, \#Bif\_Arg.2, 1)

\hypertarget{wordindex}{%
\subsection{WORDINDEX}\label{wordindex}}

WORDINDEX returns the character position in the first argument of a word
in the first argument. The second argument is the word position of that
word.

call CheckArgs `rANY rwWHOLE\textgreater0'

String Num

\#Bif Arg.1 \#Bif Arg.2

/* Find starting position */

Start = 1 Count = 0 do forever Start = verify(String,
\#Al1Blanks\textless Index ``\#A11Blanks'' \# ``\,'' \textgreater, `N',
Start) /\emph{ Find non-blank }/ if Start = 0 then return 0 /* Start is
beyond end \emph{/ Count = Count + 1 /} Words found \emph{/ if Count =
Num then leave Start = verify(String, \#Al1Blanks\textless Index
``\#Al11Blanks'' \# ``\,'' \textgreater, `M', Start + 1) /}

Find blank */

if Start = 0 then return 0 /* Start is beyond end */ end return Start

\hypertarget{wordlength}{%
\subsection{WORDLENGTH}\label{wordlength}}

WORDLENGTH returns the number of characters in a word from the first
argument. The second argument is the word position of that word.

call CheckArgs `rANY rwWHOLE\textgreater0'

return length(subword(\#Bif Arg.1, \#Bif\_Arg.2, 1))

\hypertarget{wordpos}{%
\subsection{WORDPOS}\label{wordpos}}

WORDPOS finds the leftmost occurrence in the second argument of the
sequence of words in the first argument. The result is `0' or the word
position in the second argument of the first word of the matched
sequence. Third argument is a word position for the start of the search.

call CheckArgs `rANY rANY oWHOLE\textgreater0'

Phrase = \#Bif\_Arg.1

String = \#Bif Arg.2

if \#Bif\_ArgExists.3 then Start = \#Bif\_Arg.3 else Start = 1

Phrase = space (Phrase) PhraseWords = words (Phrase) if PhraseWords = 0
then return 0 String = space (String) StringWords = words (String) do
WordNumber = Start to StringWords - PhraseWords + 1 if Phrase ==
subword(String, WordNumber, PhraseWords) then return WordNumber end
WordNumber return 0

\hypertarget{words}{%
\subsection{WORDS}\label{words}}

WORDS counts the number of words in its argument.

call CheckArgs `rANY'

do Count = 0 by 1 if subword(\#Bif Arg.1, Count + 1) == '\,' then return
Count end Count

\hypertarget{xrange}{%
\subsection{XRANGE}\label{xrange}}

XRANGE returns an ordered string of all valid character encodings in the
specified range.

call CheckArgs `oPAD oPAD'

if \#Bif\_ArgExists.1 then \#Bif\_Arg.1 mr if \#Bif\_ArgExists.2 then
\#Bif Arg.2 \#Response = Config Xrange(\#Bif Arg.1, \#Bif Arg.2) return
\#Outcome

\hypertarget{arithmetic-built-in-functions}{%
\section{Arithmetic built-in
functions}\label{arithmetic-built-in-functions}}

These functions perform arithmetic at the numeric settings current at
the invocation of the built-in function. Note that CheckArgs formats any
`NUM' (numeric) argument.

\hypertarget{abs}{%
\subsection{ABS}\label{abs}}

ABS returns the absolute value of its argument.

call CheckArgs `rNUM' Number=\#Bif Arg.1

if left (Number,1) = `-' then Number = substr (Number, 2) return Number

\hypertarget{format}{%
\subsection{FORMAT}\label{format}}

FORMAT formats its first argument. The second argument specifies the
number of characters to be used for the integer part and the third
specifies the number of characters for the decimal part. The fourth
argument specifies the number of characters for the exponent and the
fifth determines when exponeniial notation is used.

call CheckArgs, `rNUM OWHOLE\textgreater=0 OWHOLE\textgreater=0
OWHOLE\textgreater=0 OWHOLE\textgreater=0'

if \#Bif\_ArgExists.2 then Before \#Bif Arg.2 if \#Bif\_ArgExists.3 then
After \#Bif Arg.3 if \#Bif\_ArgExists.4 then Expp \#Bif Arg.4

if \#Bif\_ArgExists.5 then Expt = \#Bif Arg.5

/* In the simplest case the first is the only argument. */ Number=\#Bif
Arg.1

if \#Bif\_Arg.0 \textless{} 2 then return Number

/* Dissect the Number. It is in the normal Rexx format. */ parse var
Number Mantissa `E' Exponent

if Exponent == '\,' then Exponent = 0 Sign = 0 if left (Mantissa,1) ==
`-' then do Sign = 1 Mantissa = substr(Mantissa, 2) end parse var
Mantissa Befo `.' Afte

/* Count from the left for the decimal point. */

Point = length (Befo)

/* Sign Mantissa and Exponent now reflect the Number. Befo Afte and
Point reflect Mantissa. */

/* The fourth and fifth arguments allow for exponential notation.
\emph{/ /} Decide whether exponential form to be used, setting ShowExp.
\emph{/ ShowExp = 0 if \#Bif\_ArgExists.4 \#Bif ArgExists.5 then do if
\#Bif\_ArgExists.5 then Expt = \#Digits.\#Level /} Decide whether
exponential form to be used. \emph{/ if (Point + Exponent)
\textgreater{} Expt then ShowExp = 1 /} Digits before rule. \emph{/
LeftOfPoint = 0 if length(Befo) \textgreater{} 0 then LeftOfPoint = Befo
/} Value left of the point */

/* Digits after point rule for exponentiation: */

/* Count zeros to right of point. */

ze=O0

do while substr(Afte,z+1,1) == `0' Zeze+41tl end

if LeftOfPoint = 0 \& (z - Exponent) \textgreater{} 5 then ShowExp = 1

/* An extra rule for exponential form: */ if \#Bif\_ArgExists.4 then if
Expp = 0 then ShowExp = 0

/* Construct the exponential part of the result. */ if ShowExp then do

Exponent = Exponent + ( Point - 1 ) Point = 1 /* As required for
'SCIENTIFIC! */ if \#Form.\#Level == `ENGINEERING' then do while
Exponent//3 = 0

Point = Point+1l

Exponent = Exponent-1

end end

if \ShowExp then Point = Point + Exponent end /* Expp or Expt given
\emph{/ else do /} Even if Expp and Expt are not given, exponential
notation will be used if the original Number+0 done by CheckArgs led to
it. */ if Exponent = 0 then do ShowExp = 1

110 111

end end

/* ShowExp now indicates whether to show an exponent, Exponent is its
value. \emph{/ /} Make this a Number without a point.

Integer = Befo\textbar\textbar Afte

*/

/* Make sure Point position isn't disjoint from Integer. \emph{/ if
Point\textless1 then do /} Extra zeros on the left. */

Integer = Point = 1 end

copies(`0',1 - Point)

if Point \textgreater{} length(Integer) then Integer =
left(Integer,Point,`0') /* And maybe on the right.

/* Deal with right of decimal point first since that can affect the

\textbar\textbar{} Integer

left. Ensure the requested number of digits there. Afters =
length(Integer) -Point

if \#Bif\_ArgExists.3 = 0 then After = /* Make Afters match the
requested After */

do while Afters \textless{} After Afters = Afters+1 Integer = Integer'0'
end

if Afters \textgreater{} After then do

/* Round by adding 5 at the right place.

Afters

r=substr (Integer, Point + After + 1,

Integer = left (Integer,

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
\end{enumerate}

Point + After)

if r \textgreater= `5' then Integer = Integer + 1 /* This can leave the
result zero. If Integer = 0 then Sign = 0

/* The case when rounding makes the integer longer is an awkward

*/

*/

one. The exponent will have to be adjusted. */ if length(Integer)
\textgreater{} Point + After then do

Point = Point+l end

if ShowExp = 1 then do

Exponent=Exponent + (Point - 1)

Point = 1 /* As required for 'SCIENTIFIC! */

if form() = `ENGINEERING' then do while Exponent//3 = 0

Point = Point+

1

Exponent = Exponent-1

end end

t = Point-length (Integer) if t \textgreater{} 0 then Integer =
Integer\textbar\textbar copies(`0',t)

end /* Rounded */

/* Right part is final if After \textgreater{} 0 then Afte else Afte

now. */ `.'\textbar{} \textbar substr (Integer, Point+1,After)

/* Now deal with the integer part of the result. Integer = left
(Integer, Point)

if \#Bif ArgExists.2 =

0 then Before

/* Make Point match Before */ if Point \textgreater{} Before - Sign then
call Raise

do while Point\textless Before Point = Point+1l Integer = '0'Integer end

40.38,

*/

Point + Sign /* Note default.

2,

/* Find the Sign position and blank leading zeroes.

re ter Triggered = 0

do j = 1 to length (Integer) Digit = substr(Integer,j,1) /* Triggered is
set when sign inserted or blanking finished. if Triggered = 1 then do

r= r\textbar\textbar Daigit iterate end

*/

/* Note default.

\#Bif Arg.1

*/

*/

*/

*/

\emph{/ /} If before sign insertion point then blank out zero. */

if Digit = `0' then if ats cael = `0' \& j+l\textless length(Integer)
then do re r ' ' iterate end /* j is the sign insertion point. */ if
Digit = `0' \& j = length(Integer) then Digit = ' '

if Sign = 1 then Digit = `-' r= xr\textbar\textbar Digit Triggered = 1
end j Number = r\textbar\textbar Afte

if ShowExp = 1 then do /* Format the exponent. \emph{/ Expart = '\,'
SignExp = 0 if Exponent\textless0 then do SignExp = 1 Exponent =
-Exponent end /} Make the exponent to the requested width. */ if \#Bif\_
ArgExists.4 = 0 then Expp = length (Exponent) if length(Exponent)
\textgreater{} Expp then call Raise 40.38, 4, \#Bif\_ Arg.1
Exponent=right (Exponent,Expp,`0') if Exponent = 0 then do if \#Bif\_
ArgExists.4 then Expart = copies(' ',expp+2) end else if SignExp = 0
then Expart else Expart Number = Number \textbar{} \textbar Expart end
return Number

`E+' Exponent `E-' Exponent

\hypertarget{max}{%
\subsection{MAX}\label{max}}

MAX returns the largest of its arguments.

if \#Bif\_Arg.0 \textless1 then call Raise 40.3, 1 call CheckArgs
`rNUM'\textbar\textbar copies(' rNUM', \#Bif Arg.0 - 1)

Max = \#Bif Arg.1

do i = 2 to \#Bif Arg.0 by 1 Next = \#Bif Arg.i if Max \textless{} Next
then Max = Next end i

return Max

\hypertarget{min}{%
\subsection{MIN}\label{min}}

MIN returns the smallest of its arguments.

if \#Bif\_Arg.0 \textless1 then call Raise 40.3, 1 call CheckArgs
`rNUM'\textbar\textbar copies(' rNUM', \#Bif Arg.0 - 1)

Min = \#Bif Arg.1

do i = 2 to \#Bif Arg.0 by 1 Next = \#Bif Arg.i if Min \textgreater{}
Next then Min = Next end i

return Min

\hypertarget{sign}{%
\subsection{SIGN}\label{sign}}

SIGN returns `1', `0' or `-1' according to whether its argument is
greater than, equal to, or less than zero.

call CheckArgs `rNUM'

Number = \#Bif Arg.1

select

when Number \textless{} 0 then Output = -1 when Number = 0 then Output =
0 when Number \textgreater{} 0 then Output = 1

end return Output

\hypertarget{trunc}{%
\subsection{TRUNC}\label{trunc}}

TRUNC returns the integer part of its argument, or the integer part plus
a number of digits after the decimal point, specified by the second
argument.

call CheckArgs `rNUM oWHOLE\textgreater=0'

Number = \#Bif Arg.1 if \#Bif\_ArgExists.2 then Num

\#Bif Arg.2 else Num 0

Integer =(10**Num * Number) \%1 if Num=0 then return Integer

t=length (Integer) -Num if t\textless=0 then return
`0.'right(Integer,Num,'0') else return insert(`.',Integer,t)

\hypertarget{state-built-in-functions}{%
\section{State built-in functions}\label{state-built-in-functions}}

These functions return values from the state of the execution.

\hypertarget{address}{%
\subsection{ADDRESS}\label{address}}

ADDRESS returns the name of the environment to which commands are
currently being submitted. Optionally, under control by the argument, it
also returns information on the targets of command output and the source
of command input.

call CheckArgs `oEINO'

if \#Bif\_ArgExists.1 then Optionl = \#Bif\_Arg.1 else Optionl=`N'

if Optionl == `N' then return \#Env\_Name.ACTIVE. \#Level

Tail = Optionl' .ACTIVE. '\#Level return \#Env\_Position.Tail
\#Env\_Type.Tail \#Env\_Resource.Tail

\hypertarget{arg}{%
\subsection{ARG}\label{arg}}

ARG returns information about the argument strings to a program or
routine, or the value of one of those strings.

ArgData = `OWHOLE\textgreater0 oENO' if \#Bif\_ ArgExists.2 then ArgData
= `rWHOLE\textgreater0 rENO' call CheckArgs ArgData

if \#Bif\_ArgExists.1 then return \#Arg.\#Level.0

ArgNum=\#Bif Arg.1

if \#Bif\_ArgExists.2 then return \#Arg.\#Level.ArgNum

if \#Bif\_Arg.2 ==`0O' then return \#ArgExists.\#Level.ArgNum else
return \#ArgExists.\#Level .ArgNum

\hypertarget{condition}{%
\subsection{CONDITION}\label{condition}}

CONDITION returns information associated with the current condition.

call CheckArgs `oCDEIS'

/* Values are null if this is not following a condition. */ if
\#Condition.\#Level == '\,' then do \#ConditionDescription.\#Level =
'\,' \#ConditionExtra.\#Level = '\,' \#ConditionInstruction.\#Level end

Option=\#Bif Arg.1

if Option==`C' then return \#Condition.\#Level

if Option==`D' then return \#ConditionDescription. \#Level if
Option==`E' then return \#ConditionExtra.\#Level

if Option==`I' then return \#ConditionInstruction. \#Level /* State is
the current state. */

if \#Condition.\#Level = '\,' then return ``\,'' return
\#Enabling.\#Condition.\#Level \#\#\# DIGITS

DIGITS returns the current setting of NUMERIC DIGITS. call CheckArgs
'\,'

return \#Digits.\#Level

\hypertarget{errortext}{%
\subsection{ERRORTEXT}\label{errortext}}

ERRORTEXT returns the unexpanded text of the message which is identified
by the first argument. A second argument of `S' selects the standard
English text, otherwise the text may be translated to another national
language. This translation is not shown in the code below.

call CheckArgs `r0\_90 oSN'

msgcode = \#Bif Arg.1

if \#Bif\_ ArgExists.2 then Option else Option

return \#ErrorText .msgcode

\hypertarget{form}{%
\subsection{FORM}\label{form}}

FORM returns the current setting of NUMERIC FORM.

\#Bif\_Arg.2 tint

call CheckArgs '\,'

return \#Form.\#Level

\hypertarget{fuzz}{%
\subsection{FUZZ}\label{fuzz}}

FUZZ returns the current setting of NUMERIC FUZZ.

call CheckArgs '\,'

return \#Fuzz.\#Level

\hypertarget{sourceline}{%
\subsection{SOURCELINE}\label{sourceline}}

If there is no argument, SOURCELINE returns the number of lines in the
program, or `0' if the source program is not being shown on this
execution. If there is an argument it specifies the number of the line
of the source program to be returned.

call CheckArgs `oWHOLE\textgreater0'

if \#Bif\_ArgExists.1 then return \#SourceLine.0 Num = \#Bif\_Arg.1 if
Num \textgreater{} \#SourceLine.0O then call Raise 40.34, Num,
\#SourceLine.0 return \#SourceLine.Num

\hypertarget{trace}{%
\subsection{TRACE}\label{trace}}

TRACE returns the trace setting currently in effect, and optionally
alters the setting.

call CheckArgs `oACEFILNOR' /* Also checks for `?!' \emph{/ /} With no
argument, this a simple query. */ Output=\#Tracing.\#Level

if \#Interactive.\#Level then Output = `?'\textbar\textbar Output if
\#Bif\_ArgExists.1 then return Output

Value=\#Bif Arg.1

\#Interactive.\#Level=0

/* A question mark sets the interactive flag. */ if left(Value,1)==`?'
then do

\#Interactive.\#Level = 1

Value=substr (Value, 2)

end /* Absence of a letter leaves the setting unchanged. */ if
Value=='\,' then do

Value=translate (left (Value,1)) if Value==`0O' then
\#Interactive.\#Level=`0' \#Tracing.\#Level = Value end return Output

10.4 Conversion built-in functions

Conversions between Binary form, Decimal form, and heXadecimal form do
not depend on the encoding (see nnn) of the character data.

Conversion to Coded form gives a result which depends on the encoding.
Depending on the encoding, the result may be a string that does not
represent any sequence of characters.

\hypertarget{b2xx}{%
\subsection{B2xX}\label{b2xx}}

B2X performs binary to hexadecimal conversion.

call CheckArgs `rBIN'

String = space (\#Bif Arg.1,0) return ReRadix(String,2,16)

\hypertarget{bitand}{%
\subsection{BITAND}\label{bitand}}

The functions BITAND, BITOR and BITXOR operate on encoded character
data. Each binary digit from the encoding of the first argument is
processed in conjunction with the corresponding bit from the second
argument.

call CheckArgs `rANY oANY oPAD'

Stringl = \#Bif\_ Arg.1 if \#Bif ArgExists.2 then String2

\#Bif Arg.2 else String2 rr

/* Presence of a pad implies character strings. */ if \#Bif\_ArgExists.3
then if length(Stringl) \textgreater{} length(String2) then String2=left
(String2,length(String1l),\#Bif\_Arg.3) else
Stringl=left(Stringl,length(String2),\#Bif\_Arg.3)

/* Change to manifest bit representation. \emph{/ \#Response=Config
C2B(String1) String1l=\#Outcome \#Response=Config C2B(String2)
String2=\#0utcome /} Exchange if necessary to make shorter second. */ if
length(Stringl)\textless length(String2) then do

t=Stringl

Stringl=String2

String2=t

end

/* Operate on common length of those bit strings. */ r='\,' do j=1 to
length (String2)

bl=substr (Stringl,j,1)

b2=substr (String2,j,1)

select when \#Bif=`BITAND' then b1=b1\&b2 when \#Bif=`BITOR' then
bl=b1\textbar b2

115 when \#Bif=`BITXOR' then b1=b1\&\&b2 end r=r\textbar\textbar bl end
j rer \textbar\textbar{} right (Stringl, length (String1)
-length(String2) )

/* Convert back to encoded characters. */ return x2c (b2x(r))

\hypertarget{bitor}{%
\subsection{BITOR}\label{bitor}}

See nnn

\hypertarget{bitxor}{%
\subsection{BITXOR}\label{bitxor}}

See nnn

\hypertarget{c2d}{%
\subsection{C2D}\label{c2d}}

C2D performs coded to decimal conversion.

call CheckArgs `rANY oWHOLE\textgreater=0' if length(\#Bif Arg.1)=0 then
return 0

if \#Bif\_ArgExists.2 then do /* Size specified \emph{/ Size =
\#Bif\_Arg.2 if Size = 0 then return 0 /} Pad will normally be zeros
\emph{/ t=right (\#Bif Arg.1,Size,left (xrange(),1)) /} Convert to
manifest bit \emph{/ call Config C2B t /} And then to signed decimal.
\emph{/ Sign = Left (\#Outcome,1) \#Outcome = substr(\#Outcome, 2)
t=ReRadix (\#Outcome, 2,10) /} Sign indicates 2s-complement. \emph{/ if
Sign then t=t-2\textbf{length(\#Outcome) if abs(t) \textgreater{} 10 }
\#Digits.\#Level - 1 then call Raise 40.35, t return t end /} Size not
specified. */ call Config C2B \#Bif\_Arg.1 t = ReRadix(\#Outcome, 2,10)
if t \textgreater{} 10 ** \#Digits.\#Level - 1 then call Raise 40.35, t
return t

10.4.6 C2Xx C2X performs coded to hexadecimal conversion.

call CheckArgs `rANY' if length(\#Bif Arg.1) = 0 then return '\,'

call Config C2B \#Bif\_Arg.1 return ReRadix (\#Outcome,2,16)

\hypertarget{d2c}{%
\subsection{D2C}\label{d2c}}

D2C performs decimal to coded conversion.

`rWHOLENUM\textgreater=0'! `rWHOLENUM rWHOLE\textgreater=0'

if \#Bif\_ArgExists.2 then ArgData else ArgData call CheckArgs ArgData

/* Convert to manifest binary \emph{/ Subject = abs(\#Bif Arg.1) r =
ReRadix(Subject,10,2) /} Make length a multiple of 8, as required for
Config B2C */ Length = length(r) do while Length//8 = 0 Length =
Length+1 end r= right(r,Length,`0')

/* 2s-complement for negatives. */ if \#Bif\_Arg.1\textless0 then do
Subject = 2**length(r)-Subject r = ReRadix(Subject,10,2) end /* Convert
to characters */ \#Response = Config B2C(r) Output = \#Outcome if
\#Bif\_ArgExists.2 then return Output

/* Adjust the length with appropriate characters. */ if
\#Bif\_Arg.1\textgreater=0 then return right (Output,
\#Bif\_Arg.2,left(xrange(),1)) else return right (Output, \#Bif
Arg.2,right (xrange(),1))

\hypertarget{d2x}{%
\subsection{D2X}\label{d2x}}

D2X performs decimal to hexadecimal conversion. if \#Bif\_ArgExists.2
then ArgData = `rWHOLENUM\textgreater=0' else ArgData = `rWHOLENUM
rWHOLE\textgreater=0'

call CheckArgs ArgData

/* Convert to manifest hexadecimal \emph{/ Subject = abs(\#Bif Arg.1 ) r
= ReRadix(Subject,10,16) /} Twos-complement for negatives */ if
\#Bif\_Arg.1\textless0 then do Subject = 16**length(r) -Subject r =
ReRadix(Subject,10,16) end if \#Bif\_ArgExists.2 then return r /* Adjust
the length with appropriate characters. */ if \#Bif\_
Arg.1\textgreater=0 then return right(r,\#Bif Arg.2,`0') else return
right(r,\#Bif Arg.2,`F')

\hypertarget{x2b}{%
\subsection{X2B}\label{x2b}}

X2B performs hexadecimal to binary conversion.

call CheckArgs `rHEX'

Subject = \#Bif Arg.1

if Subject == '\,' then return '\,'

/* Blanks were checked by CheckArgs, here they are ignored. */ Subject =
space (Subject, 0)

return ReRadix(translate (Subject) ,16,2)

\hypertarget{x2c}{%
\subsection{X2C}\label{x2c}}

X2C performs hexadecimal to coded character conversion.

call CheckArgs `rHEX'

Subject = \#Bif Arg.1

if Subject == '\,' then return '\,'

Subject = space (Subject, 0)

/* Convert to manifest binary */

r = ReRadix(translate (Subject) ,16,2)

/* Convert to character */

Length = 8*( (length (Subject) +1) \%2) \#Response = Config
B2C(right(r,Length,`0')) return \#Outcome

\hypertarget{x2d}{%
\subsection{X2D}\label{x2d}}

X2D performs hexadecimal to decimal conversion.

call CheckArgs `rHEX OWHOLE\textgreater=0'

Subject = \#Bif Arg.1 if Subject == '\,' then return `0'

Subject = translate (space (Subject,0))

if \#Bif ArgExists.2 then

Subject = right (Subject, \#Bif Arg.2,`0') if Subject =='\,' then return
`0' /* Note the sign */ if \#Bif ArgExists.2 then SignBit

else SignBit

/* Convert to decimal \emph{/ r = ReRadix(Subject,16,10) /}
Twos-complement */ if SignBit then r = 2**(4*\#Bif Arg.2) - r if
abs(r)\textgreater10 ** \#Digits.\#Level - 1 then call Raise 40.35, t
return r

left (x2b (Subject) ,1) ror

\hypertarget{inputoutput-built-in-functions}{%
\section{Input/Output built-in
functions}\label{inputoutput-built-in-functions}}

The configuration shall provide the ability to access streams. Streams
are identified by character string identifiers and provide for the
reading and writing of data. They shall support the concepts of
characters, lines, and positioning. The input/output built-in functions
interact with one another, and they make use of Config\_ functions, see
nnn. When the operations are successful the following characteristics
shall be exhibited: - The CHARIN/CHAROUT functions are insensitive to
the lengths of the arguments. The data written to a stream by CHAROUT
can be read by a different number of CHARINs. - The CHARIN/CHAROUT
functions are reflective, that is, the concatenation of the data read
from a persistent stream by CHARIN (after positioning to 1, while
CHARS(Stream)\textgreater0), will be the same as the concatenation of
the data put by CHAROUT. - All characters can be used as CHARIN/CHAROUT
data. - The CHARS(Stream, `N') function will return zero only when a
subsequent read (without positioning) is guaranteed to raise the
NOTREADY condition. - The LINEIN/LINEOUT functions are sensitive to the
length of the arguments, that is, the length of a line written by
LINEOUT is the same as the length of the string returned by successful
LINEIN of the line. - Some characters, call them line-banned characters,
cannot reliably be used as data for LINEIN/LINEOUT. If these are not
used, LINEIN/LINEOUT is reflective. If they are used, the result is not
defined. The set of characters which are line-barred is a property of
the configuration. - The LINES(Stream, `N') function will return zero
only when a subsequent LINEIN (without positioning) is guaranteed to
raise the NOTREADY condition. - When a persistent stream is repositioned
and written to with CHAROUT, the previously written data is not lost,
except for the data overwritten by this latest CHAROUT. - When a
persistent stream is repositioned and written to with LINEOUT, the
previously written data is not lost, except for the data overwritten by
this latest LINEOUT, which may leave lines partially overwritten. \#\#\#
CHARIN CHARIN returns a string read from the stream named by the first
argument.

call CheckArgs `oSTREAM oOWHOLE\textgreater0 oOWHOLE\textgreater=0'

if \#Bif\_ArgExists.1 then Stream else Stream \#StreamState.Stream =
'\,' /* Argument 2 is positioning. */ if \#Bif\_ArgExists.2 then do
\#Response = Config Stream Position(Stream, `CHARIN', \#Bif Arg.2)

\#Bif Arg.1

if left(\#Response, 1) == `R' then call Raise 40.41, 2, \#Bif\_ Arg.2 if
left(\#Response, 1) == `T' then call Raise 40.42,Stream end

/* Argument 3 is how many. \emph{/ if \#Bif\_ArgExists.3 then Count else
Count if Count = 0 then do call Config Stream Charin Stream, `NULL' /}
``Touch'' the stream \emph{/ return '! end /} The unit may be eight bits
(as characters) or one character. */ call Config Stream Query Stream

\#Bif Arg.3 1

Mode = \#Outcome

do until Count = 0

\#Response = Config Stream Charin(Stream, `CHARIN')

if left(\#Response, 1) == `N' then do if left (\#Response, 1) == `E'
then \#StreamState.Stream = `ERROR' /* This call will return. */ call
\#Raise `NOTREADY', Stream, substr(\#Response, 2) leave end

r = r\textbar{} \textbar\#Outcome

Count = Count-1

end if Mode == `B' then do call Config B2C r r = \#Outcome end return r

\hypertarget{charout}{%
\subsection{CHAROUT}\label{charout}}

CHAROUT returns the count of characters remaining after attempting to
write the second argument to the stream named by the first argument.

call CheckArgs `oSTREAM oANY oWHOLE\textgreater0'

if \#Bif\_ArgExists.1 then Stream else Stream

\#Bif Arg.1

\#StreamState.Stream = '\,' if \#Bif ArgExists.2 \& \#Bif\_ArgExists.3
then do /* Position to end of stream. */ \#Response = Config Stream
Close (Stream) if left (\#Response,1) == `T' then call Raise
40.42,Stream return 0 end

if \#Bif\_ArgExists.3 then do /* Explicit positioning. */ \#Response =
Config Stream Position(Stream, `CHAROUT', \#Bif Arg.3)

if left (\#Response,1) == `T' then call Raise 40.42,Stream if
left(\#Response, 1) == `R' then call Raise 40.41, 3, \#Bif\_ Arg.3 end

if \#Bif\_ArgExists.2 \textbar{} \#Bif\_Arg.2 == '\,' then do call
Config Stream \_Charout Stream, `NULL' /* ``Touch'' the stream */ return
0 end

String = \#Bif Arg.2 call Config Stream Query Stream Mode = \#Outcome if
Mode == `B' then do call Config C2B String String = \#Outcome Stride = 8
Residue = length(String)/8 end else do Stride = 1 Residue = length
(String) end

Cursor = 1 do while Residue\textgreater0 Piece = substr (String, Cursor,
Stride) Cursor = Cursor+Stride call Config Stream Charout Stream, Piece
if left (\#Response, 1) == `N' then do if left (\#Response, 1) == `E'
then \#StreamState.Stream = `ERROR' call \#Raise `NOTREADY', Stream,
substr(\#Response, 2)

return Residue end Residue = Residue - 1 end return 0

\hypertarget{chars}{%
\subsection{CHARS}\label{chars}}

CHARS indicates whether there are characters remaining in the named
stream. Optionally, it returns a

count of the characters remaining and immediately available.

call CheckArgs `oSTREAM oCN'

if \#Bif\_ArgExists.1 then Stream = \#Bif\_Arg.1 else Stream = '\,'

if \#Bif\_ArgExists.2 then Option = \#Bif Arg.2 else Option = `N'

call Config Stream Count Stream, `CHARS', Option return \#Outcome

\hypertarget{linein}{%
\subsection{LINEIN}\label{linein}}

LINEIN reads a line from the stream named by the first argument, unless
the third argument is Zero.

call CheckArgs `oSTREAM oOWHOLE\textgreater0 oOWHOLE\textgreater=0'

if \#Bif\_ArgExists.1 then Stream else Stream \#StreamState.Stream =
'\,' if \#Bif\_ArgExists.2 then do \#Response = Config Stream
Position(Stream, `LINEIN', \#Bif Arg2) if left (\#Response, 1) `T' then
call Raise 40.42,Stream if left (\#Response, 1) `R' then call Raise
40.41, 2, \#Bif\_ Arg.2 end if \#Bif\_ArgExists.3 then Count \#Bif Arg.3
else Count 1 if Count\textgreater1 then call Raise 40.39, Count if Count
= 0 then do call Config Stream Charin Stream, `NULL' /* ``Touch'' the
stream \emph{/ return `! end /* A configuration may recognise lines even
in 'binary' mode. }/ call Config Stream Query Stream Mode = \#Outcome re
ter t = \#Linein Position.Stream /* Config Stream Charin will alter
\#Linein Position. */ do until t = \#Linein Position.Stream \#Response =
Config Stream Charin(Stream, `LINEIN') if left(\#Response, 1) == `N'
then do if left (\#Response, 1) == `E' then \#StreamState.Stream =
`ERROR' call \#Raise `NOTREADY', Stream, substr(\#Response, 2) leave end
r = r\textbar\textbar\#Outcome end if Mode == `B' then do call Config
B2C r r= \#Outcome end return r

\#Bif Arg.1

\hypertarget{lineout}{%
\subsection{LINEOUT}\label{lineout}}

LINEOUT returns `1' or `0', indicating whether the second argument has
been successfully written to the

stream named by the first argument. A result of `1' means an
unsuccessful write. call CheckArgs `oSTREAM oANY oWHOLE\textgreater0'

if \#Bif\_ArgExists.1 then Stream else Stream

\#Bif Arg.1

\#StreamState.Stream = '\,'

if \#Bif ArgExists.2 \& \#Bif\_ArgExists.3 then do

/* Position to end of stream.

*/

\#Response = Config Stream Close (Stream) if left (\#Response,1) == `T'
then call Raise 40.42,Stream

return 0 end

if \#Bif\_ArgExists.3 then do

\#Response = Config Stream Position(Stream, `LINEOUT', \#Bif Arg.3)

if left(\#Response, 1) == `T' then call Raise 40.42,Stream if left
(\#Response, 1) == `R' then call Raise 40.41, 3, \#Bif\_Arg.3 end

if \#Bif ArgExists.2 then do

call Config Stream \_Charout Stream, '\,' /* ``Touch'' the stream */

return 0 end

String \#Bif Arg.2 Stride 1 call Config Stream Query Stream Mode =
\#Outcome if Mode == `B' then do call Config C2B String String =
\#Outcome Stride = 8 Residue = length(String)/8 end else do Stride = 1
Residue = length(String) end Cursor = 1 do while Residue \textgreater{}
0

Piece = substr (String, Cursor, Stride)

Cursor = Cursor+Stride

call Config Stream Charout Stream, Piece then do

then \#StreamState.Stream = `ERROR'

call \#Raise `NOTREADY', Stream, substr(\#Response, 2)

if left(\#Response, 1) == `N' if left (\#Response, 1) == `E' return 1
end Residue = Residue-1 end call Config Stream Charout Stream, return 0
\#\#\# LINES

`EOL'

LINES returns the number of lines remaining in the named stream.

call CheckArgs `oSTREAM oCN'

if \#Bif\_ArgExists.1 then Stream else Stream if \#Bif\_ ArgExists.2
then Option else Option

Call Config Stream Count Stream, return \#Outcome

\hypertarget{qualify}{%
\subsection{QUALIFY}\label{qualify}}

\#Bif Arg.1 ter \#Bif\_Arg.2 tint

LINES', Option

QUALIFY returns a name for the stream named by the argument. The two
names are currently associated with the same resource and the result of
QUALIFY may be more persistently associated with

that resource. call CheckArgs `oSTREAM'

if \#Bif\_ArgExists.1 then Stream else Stream

\#Bif Arg.1 \#Response = Config Stream Qualified (Stream) return
\#Outcome

\hypertarget{stream}{%
\subsection{STREAM}\label{stream}}

STREAM returns a description of the state of, or the result of an
operation upon, the stream named by the first argument.

/* Third argument is only correct with `C' */

if \#Bif ArgExists.2 \& translate(left(\#Bif Arg.2, 1)) == `C' then
ArgData = `rSTREAM rCDS rANY'

else ArgData = `rSTREAM oCDS'

call CheckArgs ArgData

Stream = \#Bif\_Arg.1

if \#Bif\_ ArgExists.2 then Operation = \#Bif\_Arg.2 else Operation =
`S' Select when Operation == `C' then do

call Config Stream Command Stream, \#Bif Arg.3 return \#Outcome

end when Operation == `D' then do \#Response = Config Stream State
(Stream) return substr(\#Response, 2) end when Operation == `S' then do
if StreamState.Stream == `ERROR' then return `ERROR' \#Response = Config
Stream State (Stream)

`N' then return `READY' `U' then return `UNKNOWN'

if left (\#Response, if left (\#Response, return `NOTREADY' end

end

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  ==
\item
  ==
\end{enumerate}

\hypertarget{other-built-in-functions}{%
\section{Other built-in functions}\label{other-built-in-functions}}

\hypertarget{date}{%
\subsection{DATE}\label{date}}

DATE with fewer than two arguments returns the local date. Otherwise it
converts the second argument (which has a format given by the third
argument) to the format specified by the first argument. If there are
fourth or fifth arguments, they describe the treatment of separators
between fields of the date.

call CheckArgs `oBDEMNOSUW oANY oOBDENOSU oSEP oSEP' /* If the third
argument is given then the second is mandatory. */ if \#Bif\_ArgExists.3
\& \#Bif\_ArgExists.2 then

call Raise 40.19, '\,', \#Bif Arg.3

if \#Bif\_ ArgExists.1 then Option else Option

\#Bif Arg.1 tint

/* The date/time is `frozen' throughout a clause. \emph{/ if
\#ClauseTime.\#Level == '\,' then do \#Response = Config Time ()
\#ClauseTime.\#Level = \#Time \#ClauseLocal.\#Level = \#Time +
\#Adjust\textless Index ``\#Adjust'' \# ``\,'' \textgreater{} end /}
English spellings are used, even if messages not in English are used. */
Months = `January February March April May June July', `August September
October November December' WeekDays = `Monday Tuesday Wednesday Thursday
Friday Saturday Sunday'

/* If there is no second argument, the current date is returned. */ if
\#Bif ArgExists.2 then return DateFormat (\#ClauseLocal.\#Level, Option)

/* If there is a second argument it provides the date to be converted.
*/

Value = \#Bif\_Arg.2 if \#Bif\_ ArgExists.3 then InOption else InOption
if Option == `S' then OutSeparator else OutSeparator if
\#Bif\_ArgExists.4 then do

if OutSeparator == `x' then call OutSeparator = \#Bif.Arg.4 end

if InOption == `S' then InSeparator

else InSeparator if \#Bif\_ArgExists.5 then do

\#Bif\_Arg.3 'nt

translate (Option, ``xx/x //x'',``BDEMNOUW'')

Raise 40.46, Option, 4

translate(InOption,``xx/ //'',``BDENOU'')

if InSeparator == `x' then call Raise 40.46, InOption, 5

InSeparator = \#Bif.Arg.5 end /* First try for Year Month Day */ Logic =
`NS' select when InOption == `N' then do if InSeparator == '\,' then do

if length(Value)\textless9 then return Year = right (Value, 4)

MonthIs = substr (right (Value,7),1,3) Day = left (Value, length (Value)
-7)

end else

parse var Value Day (InSeparator) MonthIs (InSeparator) Year

do Month = 1 to 12 if left(word(Months, Month), 3)

== MonthIs then leave

parse var Value Year (InSeparator) Month (InSeparator) Day

end Month

end

when InOption == `S' then if InSeparator == '\,' then

parse var Value Year +4 Month +2 Day

else

otherwise Logic = `EOU' /* or BD */

end

/* Next try for year without century */

parse var Value Day (InSeparator) Month (InSeparator) YY

parse var Value YY (InSeparator) Month (InSeparator) Day

parse var Value Month (InSeparator) Day (InSeparator) YY

if logic = `EOU' then Select when InOption == `E' then if InSeparator ==
'\,' then parse var Value Day +2 Month +2 YY else when InOption == `O'
then if InSeparator == '\,' then parse var Value YY +2 Month +2 Day else
when InOption == `U' then if InSeparator == '\,' then parse var Value
Month +2 Day +2 YY else otherwise Logic = `BD' end if Logic = `EOU' then
do

/* The century is assumed, on the basis of the current year. */

if datatype(YY,`W')=0 then return

YearNow = left(`DATE'(`S'),4)

Year = YY

do while Year \textless{} YearNow-50 Year = Year + 100

end

end /* Century assumption */

if Logic \textless\textgreater{} `BD' then do /* Convert Month \& Day to
Days of year. \emph{/ if datatype(Month,`W')=0 \textbar{}
datatype(Day,`W')=0 \textbar{} datatype(Year,`W')=0 then return Days =
word(`0 31 59 90 120 151 181 212 243 273 304 334',Month), +
(Month\textgreater2)}Leap(Year) + Day-1

end else

if datatype(Value,`W')=0 then

return

if InOption == `D' then do

Year = left(`DATE' (`S'),4)

Days = Value - 1 /* `D' includes current day */ end

/* Convert to BaseDays */ if InOption \textless\textgreater{} `B' then

BaseDays = (Year-1)*365 + (Year-1)\%4 - (Year-1)\%100 + (Year-1)\%400 +
Days else

Basedays = Value

/* Convert to microseconds from 0001 \emph{/ Micro = BaseDays } 86400 *
1000000

/* Reconvert to check the original. (eg for Month = 99) */

if DateFormat (Micro,InOption,InSeparator) == Value then call Raise
40.19, Value, InOption

return DateFormat (Micro, Option, OutSeparator)

DateFormat: /* Convert from microseconds to given format. */ parse value
Time2Date(arg(1)) with, Year Month Day Hour Minute Second Microsecond
Base Days

select

when arg(2) == `B' then

return Base when arg(2) == `D' then

return Days when arg(2) == `E' then

return right (Day,2,`0') (arg(3)) right (Month,2,`0') (arg(3)) right
(Year,2,`0') when arg(2) == `M' then

return word (Months ,Month) when arg(2) == `N' then

return (Day) (arg(3)) left (word(Months,Month),3) (arg(3))right
(Year,4,`0') when arg(2) == `O' then

return right (Year,2,`0') (arg(3))right (Month,2,`0') (arg(3))right
(Day,2,`0') when arg(2) == `S' then

return right (Year,4,`0') (arg(3))right (Month,2,`0') (arg(3))right
(Day,2,`0') when arg(2) == `U' then

return right (Month,2,`0') (arg(3)) right (Day,2,`0') (arg(3)) right
(Year,2,`0') otherwise /* arg(2) == `W' */

return word (Weekdays, 1+Base//7)

end

\hypertarget{queued}{%
\subsection{QUEUED}\label{queued}}

QUEUED returns the number of lines remaining in the external data queue.

call CheckArgs '\,'

\#Response = Config Queued() return \#Outcome

\hypertarget{random}{%
\subsection{RANDOM}\label{random}}

RANDOM returns a quasi-random number.

call CheckArgs `oWHOLE\textgreater=0 oWHOLE\textgreater=0
oWHOLE\textgreater=0'

if \#Bif\_Arg.0 = 1 then do Minimum = 0

Maximum = \#Bif Arg.1 if Maximum\textgreater100000 then call Raise
40.31, Maximum

end else do if \#Bif\_ArgExists.1 then Minimum = \#Bif Arg.1 else
Minimum = 0 if \#Bif\_ArgExists.2 then Maximum = \#Bif\_Arg.2 else
Maximum = 999

end

if Maximum-Minimum\textgreater100000 then call Raise 40.32, Minimum,
Maximum

if Maximum-Minimum\textless0 then call Raise 40.33, Minimum, Maximum

if \#Bif\_ArgExists.3 then call Config Random Seed \#Bif\_Arg.3 call
Config Random Next Minimum, Maximum return \#Outcome

\hypertarget{symbol}{%
\subsection{SYMBOL}\label{symbol}}

The function SYMBOL takes one argument, which is evaluated. Let String
be the value of that argument. If Config\_Length(String) returns an
indicator `E' then the SYNTAX condition 23.1 shall be raised. Otherwise,
if the syntactic recognition described in section nnn would not
recognize String as a symbol then the result of the function SYMBOL is
'BAD''.

If String would be recognized as a symbol the result of the function
SYMBOL depends on the outcome of accessing the value of that symbol, see
nnn. If the final use of Var\_Value leaves the indicator with value `D'
then the result of the function SYMBOL is `LIT', otherwise `VAR'.

\hypertarget{time}{%
\subsection{TIME}\label{time}}

TIME with less than two arguments returns the local time within the day,
or an elapsed time. Otherwise it converts the second argument (which has
a format given by the third argument) to the format specified by the
first argument.

call CheckArgs `oCEHLMNORS oANY oCHLMNS' /* If the third argument is
given then the second is mandatory. */ if \#Bif\_ArgExists.3 \&
\#Bif\_ArgExists.2 then

call Raise 40.19, '\,', \#Bif Arg.3

if \#Bif\_ ArgExists.1 then Option else Option

\#Bif Arg.1 tint

/* The date/time is `frozen' throughout a clause. */

if \#ClauseTime.\#Level == '\,' then do \#Response = Config Time ()
\#ClauseTime.\#Level = \#Time \#ClauseLocal.\#Level = \#Time +
\#Adjust\textless Index ``\#Adjust'' \# ``\,'' \textgreater{} end

/* If there is no second argument, the current time is returned. */ if
\#Bif ArgExists.2 then return TimeFormat (\#ClauseLocal.\#Level, Option)

/* If there is a second argument it provides the time to be converted.
*/ if pos(Option, `ERO') \textgreater{} 0 then call Raise 40.29, Option
InValue = \#Bif Arg.2

if \#Bif\_ArgExists.3 then InOption = \#Bif\_Arg.3 else InOption =

HH = 0

MM = 0

ss = 0

HourAdjust = 0

select

when InOption == `C' then do parse var InValue HH `:' . +1 MM +2 XX

if HH = 12 then

125 HH = 0

if XX == `pm' then HourAdjust = 12 end when InOption == `H' then HH =
InValue when InOption == `L' \textbar{} InOption == `N' then parse var
InValue HH `:' MM `:' SS when InOption == `M' then MM = InValue
otherwise SS = InValue

end

if datatype(HH,`W')=0 \textbar{} datatype(MM,`W')=0 \textbar{}
datatype(SS,`N')=0 then call Raise 40.19, InValue, InOption

HH = HH + HourAdjust

/* Convert to microseconds */

Micro = trunc((((HH * 60) + MM) * 60 + SS) * 1000000)

/* There is no special message for time-out-of-range; the bad-format

message is used. */

if Micro\textless0 \textbar{} Micro \textgreater{} 24\emph{3600}1000000
then call Raise 40.19, InValue, InOption

/* Reconvert to check the original. (eg for hour = 99) */

if TimeFormat (Micro,InOption) == InValue then call Raise 40.19,
InValue, InOption

return TimeFormat (Micro, Option)

end /* Conversion */

TimeFormat: procedure /* Convert from microseconds to given format.
\emph{/ /} The day will be irrelevant; actually it will be the first day
possible. \emph{/ x = Time2Date2(arg(1)) parse value x with Year Month
Day Hour Minute Second Microsecond Base Days select when arg(2) == `C'
then select when Hour\textgreater12 then return Hour-12':`right
(Minute,2,'0')`pm' when Hour=12 then return `12:'right(Minute,2,'0')`pm'
when Hour\textgreater0 then return Hour':`right (Minute,2,'0')`am' when
Hour=0 then return `12:'right(Minute,2,'0')`am' end when arg(2) == `H'
then return Hour when arg(2) == `L' then return right
(Hour,2,`0')`:'right (Minute,2,'0')`:'right(Second,2,'0'),
\textbar\textbar{} `.'right (Microsecond,6,'0') when arg(2) == `M' then
return 60}Hour+Minute when arg(2) == `N' then return right
(Hour,2,`0')`:'right (Minute,2,'0')`:'right(Second,2,'0') otherwise /*
arg(2) == `S' \emph{/ return 3600}Hour+60*Minute+Second end

Time2Date: /* These are checks on the range of the date. */ if arg(1)
\textless{} 0 then call Raise 40.19, InValue, InOption if arg(1)
\textgreater= 315537897600000000 then call Raise 40.19, InValue,
InOption return Time2Date2 (arg(1))

10.6.1 VALUE VALUE returns the value of the symbol named by the first
argument, and optionally assigns it a new

value.

`rANY oOANY oANY'

if \#Bif ArgExists.3 then ArgData `rSYM oANY oOANY'

else ArgData call CheckArgs ArgData

Subject = \#Bif Arg.1

if \#Bif\_ArgExists.3 then do /* An external pool, or the reserved pool.
\emph{/ /} The reserved pool uses a null string as its pool identifier.
*/ Pool = \#Bif\_Arg.3

if Pool == '\,' then do Subject = `.' \textbar\textbar{} translate
(Subject) /* The dot on the name is implied. \emph{/ Value =
.environment {[}Subject{]} /} Was the translate redundant? */

if \#Bif ArgExists.2 then .environment {[}Subject{]} = \#Bif\_ Arg.2
return Value end

/* Fetch the original value */

\#Response = Config Get (Pool, Subject)

\#Indicator = left (\#Response,1)

if \#Indicator == `F' then call Raise 40.36, Subject if \#Indicator ==
`P' then

call Raise 40.37, Pool Value = \#Outcome if \#Bif\_ArgExists.2 then do
/* Set the new value. \emph{/ \#Response = Config Set(Pool,Subject,\#Bif
Arg.2) if \#Indicator == `P' then call Raise 40.37, Pool if \#Indicator
== `F' then call Raise 40.36, Subject end /} Return the original value.
\emph{/ return Value end /} Not external \emph{/ Subject =
translate(Subject) /} See nnn \emph{/ Pp = pos(Subject, `.') if p = 0
\textbar{} p = length(Subject) then do /} Not compound \emph{/
\#Response = Var Value(\#Pool, Subject, `0') /} The caller, in the code
of the standard, may need to test whether the Subject was dropped.
\emph{/ \#Indicator = left(\#Response, 1) Value = \#Outcome if \#Bif\_
ArgExists.2 then \#Response = Var Set(\#Pool, Subject, `0', \#Bif\_
Arg.2) return Value end /} Compound \emph{/ Expanded = left(Subject,p-1)
/} The stem \emph{/ do forever Start = p+l1 Pp = pos(Subject,`.',Start)
if p = 0 then p = length(Subject) Item = substr(Subject,Start,p-Start)
/} Tail component symbol \emph{/ if Item=='\,' then if
pos(left(Item,1),`0123456789') = 0 then do \#Response = Var
Value(\#Pool, Item, `0') Item = \#Outcome end /} Add tail component. */
Expanded = Expanded'.'Item end \#Response = Var Value(\#Pool, Expanded,
`1') \#Indicator = left(\#Response, 1) Value = \#Outcome if \#Bif\_
ArgExists.2 then \#Response = Var Set(\#Pool, Expanded, `1', \#Bif\_
Arg.2)

return Value

\hypertarget{queued-1}{%
\subsection{QUEUED}\label{queued-1}}

QUEUED returns the number of lines remaining in the external data queue.

call CheckArgs '\,' \#Response = Config Queued() return \#Outcome

\hypertarget{random-1}{%
\subsection{RANDOM}\label{random-1}}

RANDOM returns a quasi-random number.

call CheckArgs `oWHOLE\textgreater=0 oWHOLE\textgreater=0
oWHOLE\textgreater=0'

if \#Bif\_Arg.0 = 1 then do Minimum = 0 Maximum = \#Bif\_Arg.1 if
Maximum\textgreater100000 then call Raise 40.31, Maximum

end else do if \#Bif\_ArgExists.1 then Minimum = \#Bif Arg.1 else
Minimum = 0 if \#Bif\_ArgExists.2 then Maximum = \#Bif\_Arg.2 else
Maximum = 999

end

if Maximum-Minimum\textgreater100000 then call Raise 40.32, Minimum,
Maximum

if Maximum-Minimum\textless0 then call Raise 40.33, Minimum, Maximum

if \#Bif\_ArgExists.3 then call Config Random Seed \#Bif\_Arg.3 call
Config Random Next Minimum, Maximum return \#Outcome

\hypertarget{symbol-1}{%
\subsection{SYMBOL}\label{symbol-1}}

The function SYMBOL takes one argument, which is evaluated. Let String
be the value of that argument. If Config\_Length(String) returns an
indicator `E' then the SYNTAX condition 23.1 shall be raised. Otherwise,
if the syntactic recognition described in section nnn would not
recognize String as a symbol then the result of the function SYMBOL is
'BAD''.

If String would be recognized as a symbol the result of the function
SYMBOL depends on the outcome of accessing the value of that symbol, see
nnn. If the final use of Var\_Value leaves the indicator with value `D'
then the result of the function SYMBOL is `LIT', otherwise `VAR'.

\hypertarget{time-1}{%
\subsection{TIME}\label{time-1}}

TIME with less than two arguments returns the local time within the day,
or an elapsed time. Otherwise it converts the second argument (which has
a format given by the third argument) to the format specified by the
first argument.

call CheckArgs `oCEHLMNORS oANY oCHLMNS' /* If the third argument is
given then the second is mandatory. */ if \#Bif\_ArgExists.3 \&
\#Bif\_ArgExists.2 then

call Raise 40.19, '\,', \#Bif Arg.3

if \#Bif\_ ArgExists.1 then Option else Option

\#Bif Arg.1 tint

/* The date/time is `frozen' throughout a clause. */

if \#ClauseTime.\#Level == '\,' then do \#Response = Config Time ()
\#ClauseTime.\#Level = \#Time \#ClauseLocal.\#Level = \#Time +
\#Adjust\textless Index ``\#Adjust'' \# ``\,'' \textgreater{} end

/* If there is no second argument, the current time is returned. */ if
\#Bif ArgExists.2 then return TimeFormat (\#ClauseLocal.\#Level, Option)

/* If there is a second argument it provides the time to be

converted. */ if pos(Option, `ERO') \textgreater{} 0 then

128 call Raise 40.29, Option InValue = \#Bif Arg.2

if \#Bif\_ArgExists.3 then InOption = \#Bif\_Arg.3 else InOption = `N'
HH = 0 MM = 0 ss = 0 HourAdjust = 0 select when InOption == `C' then do
parse var InValue HH `:' . +1 MM +2 XX if XX == `pm' then HourAdjust =
12 end when InOption == `H' then HH = InValue when InOption == `L'
\textbar{} InOption == `N' then parse var InValue HH `:' MM `:' SS when
InOption == `M' then MM = InValue otherwise SS = InValue end

if \datatype(HH,`W') \textbar{} \datatype(MM,`W') \textbar{}
\datatype(SS,`N') then call Raise 40.19, InValue, InOption

HH = HH + HourAdjust

/* Convert to microseconds */

Micro = trunc((((HH * 60) + MM) * 60 + SS) * 1000000)

/* Reconvert to check the original. (eg for hour = 99) */

if TimeFormat (Micro,InOption) == InValue then call Raise 40.19,
InValue, InOption

return TimeFormat (Micro, Option)

end /* Conversion */

TimeFormat: /* Convert from microseconds to given format. */ parse value
Time2Date(arg(1)) with, Year Month Day Hour Minute Second Microsecond
Base Days select when arg(2) == `C' then if Hour\textgreater12 then
return Hour-12':`right (Minute,2,'0')`pm'

else return Hour':`right (Minute,2,'0')`am' when arg(2) == `E'
\textbar{} arg(2) == `R' then do

/* Special case first time */

if \#StartTime.\#Level == '\,' then do \#StartTime.\#Level
\#ClauseTime.\#Level return `0O'

end Output = \#ClauseTime.\#Level-\#StartTime. \#Level if arg(2) == `R'
then

\#StartTime.\#Level = \#ClauseTime.\#Level return Output * 1E-6 end /* E
or R */ when arg(2) == `H' then return Hour when arg(2) == `L' then
return right (Hour,2,`0')`:'right (Minute,2,'0')`:'right(Second,2,'0'),
\textbar\textbar{} `.'right (Microsecond,6,'0')

when arg(2) == `M' then return 60\emph{Hour+Minute when arg(2) == `N'
then return right (Hour,2,`0')`:'right
(Minute,2,'0')`:'right(Second,2,'0') when arg(2) == `O' then return
trunc(\#ClauseLocal.\#Level - \#ClauseTime.\#Level) otherwise /} arg(2)
== `S' \emph{/ return 3600}Hour+60*Minute+Second end 10.6.5 VALUE

VALUE returns the value of the symbol named by the first argument, and
optionally assigns it a new value.

`rANY oOANY oANY' `rSYM OANY oANY'

if \#Bif ArgExists.3 then ArgData else ArgData call CheckArgs ArgData

129 130

Subject = \#Bif Arg.1

if \#Bif\_ArgExists.3 then do /* An external pool, or the reserved pool.
\emph{/ /} The reserved pool uses a null string as its pool identifier.
*/ Pool = \#Bif\_Arg.3

if Pool == '\,' then do Subject = `.' \textbar\textbar{} translate
(Subject) /* The dot on the name is implied. \emph{/ Value =
.environment {[}Subject{]} /} Was the translate redundant? */

if \#Bif ArgExists.2 then .environment {[}Subject{]} = \#Bif\_ Arg.2
return Value end

/* Fetch the original value */

\#Response = Config Get (Pool, Subject)

\#Indicator = left (\#Response,1)

if \#Indicator == `F' then call Raise 40.36, Subject if \#Indicator ==
`P' then

call Raise 40.37, Pool Value = \#Outcome if \#Bif\_ArgExists.2 then do
/* Set the new value. \emph{/ \#Response = Config Set(Pool,Subject,\#Bif
Arg.2) if \#Indicator == `P' then call Raise 40.37, Pool if \#Indicator
== `F' then call Raise 40.36, Subject end /} Return the original value.
\emph{/ return Value end /} Not external \emph{/ Subject =
translate(Subject) /} See nnn \emph{/ Pp = pos(Subject, `.') if p = 0
\textbar{} p = length(Subject) then do /} Not compound \emph{/
\#Response = Var Value(\#Pool, Subject, `0') /} The caller, in the code
of the standard, may need to test whether the Subject was dropped.
\emph{/ \#Indicator = left(\#Response, 1) Value = \#Outcome if \#Bif\_
ArgExists.2 then \#Response = Var Set(\#Pool, Subject, `0', \#Bif\_
Arg.2) return Value end /} Compound \emph{/ Expanded = left(Subject,p-1)
/} The stem \emph{/ do forever Start = p+l1 Pp = pos(Subject,`.',Start)
if p = 0 then p = length(Subject) Item = substr(Subject,Start,p-Start)
/} Tail component symbol \emph{/ if Item=='\,' then if
pos(left(Item,1),`0123456789') = 0 then do \#Response = Var
Value(\#Pool, Item, `0') Item = \#Outcome end /} Add tail component. */
Expanded = Expanded'.'Item end \#Response = Var Value(\#Pool, Expanded,
`1') \#Indicator = left(\#Response, 1) Value = \#Outcome if \#Bif\_
ArgExists.2 then \#Response = Var Set(\#Pool, Expanded, `1', \#Bif\_
Arg.2)

return Value