### 2024-02-16 ARB/standards/work-in-progress/document/tex/standard/incompatibilities.tex

%preprocessed texin
\chapter{Incompatibilities}\label{incompatibilities}

\chapter{The incompatibilities from Classic Rexx
are:}\label{the-incompatibilities-from-classic-rexx-are}

Assignment of compound variables, as in ABC. = PQR., is an assignment of
references so that ABC. subsequently refers to the same object as PQR.,
as opposed to making the default value of ABC. that of PQR.. This change
was necessary to fit compound variables into the object framework, in
particular allowing USE ARG to handle compound variables as by-reference
parameters. The first reference of Annex B discouraged use of this
construct in Classic Rexx programs. ``Breakage'' of programs due to this
incompatibility is rare.

Also something in condition handling that I don't know the reason for.

\section{Call}\label{call}

The call instruction has been extended to allow for a computed name of
the callee. Syntax considerations prevent a similar thing being done for
functions.

\section{Concurrency}\label{concurrency}

Meet 17 minutes

\section{Guard}\label{guard}

Meet 17 minutes