### 2024-02-16 ARB/standards/work-in-progress/document/tex/standard/scope_purpose_and_application.tex

%preprocessed texin
\chapter{Scope, purpose, and
application}\label{scope-purpose-and-application}

\section{Scope}\label{scope}

This standard specifies the semantics and syntax of the programming
language Rexx by specifying requirements for a conforming language
processor. The scope of this standard includes

\begin{itemize}
\item
  the syntax and constraints of the Rexx language;
\item
  the semantic rules for interpreting Rexx programs;
\item
  the restrictions and limitations that a conforming language processor
  may impose;
\item
  the semantics of configuration interfaces.
\end{itemize}

This standard does not specify

\begin{itemize}
\item
  the mechanism by which Rexx programs are transformed for use by a
  data-processing system;
\item
  the mechanism by which Rexx programs are invoked for use by a
  data-processing system;
\item
  the mechanism by which input data are transformed for use by a Rexx
  program;
\item
  the mechanism by which output data are transformed after being
  produced by a Rexx program;
\item
  the encoding of Rexx programs;
\item
  the encoding of data to be processed by Rexx programs;
\item
  the encoding of output produced by Rexx programs;
\item
  the size or complexity of a program and its data that will exceed the
  capacity of any specific data-processing system or the capacity of a
  particular language processor;
\item
  all minimal requirements of a data-processing system that is capable
  of supporting a conforming language processor;
\item
  the syntax of the configuration interfaces.
\end{itemize}

\section{Purpose}\label{purpose}

The purpose of this standard is to facilitate portability of Rexx
programs for use on a wide variety of configurations.

\section{Application}\label{application}

This standard is applicable to Rexx language processors.

\section{Recommendation}\label{recommendation}

It is recommended that before detailed reading of this standard, a
reader should first be familiar with the Rexx language, for example
through reading one of the books about Rexx. It is also recommended that
the annexes should be read in conjunction with this standard.